Syntax and Semantics of Cedille

### 2020-03-03 docs/semantics/paper.tex
Fixed typo in typing rules
This fixes #140
docs: semantics: fix erasure of implicit abstraction

    \infer{\Gamma\vdash \beta\{t'\} \tpcheck \{ t \simeq t \}}{\Gamma\vdash \textit{FV}(t)\subseteq \textit{dom}(\Gamma)}  \\ \\    
    \infer{\Gamma\vdash \delta\ \mhyph\ t \tpcheck T}{\Gamma\vdash t\tpcheck \{ \absu{\lambda}{x}{\absu{\lambda}{y}{x}} \simeq \absu{\lambda}{x}{\absu{\lambda}{y}{y}}\}}  &
    \infer{\Gamma\vdash \rho\ t'\ \mhyph\ t \Leftrightarrow [t_2/x]T}
          {\Gamma\vdash t' \tpsynth t_1 \simeq t_2 & \Gamma \vdash t \Leftrightarrow [t_1/x]T} \\ \\
    \infer{\Gamma\vdash \chi\ T\ \mhyph\ t \tpcheck T'}
          {\Gamma\vdash T\tpcheck \star & \Gamma\vdash t \tpcheck T & T \cong T'} &
    \infer{\Gamma\vdash \chi\ T\ \mhyph\ t \tpsynth T}
          {\Gamma\vdash T\tpcheck \star & \Gamma\vdash t \tpsynth T' & T \cong T'} \\ \\
    \infer{\Gamma\vdash \phi\ t\ \mhyph\ t'\{t''\} \Leftrightarrow T}
          {\Gamma\vdash t\tpsynth \{t'\simeq t''\} & \Gamma\vdash t' \Leftrightarrow T}  & \
  \end{array}
  \]
\caption{Rules for checking a term against a well-kinded type ($\Gamma \vdash t \tpcheck T$)
           and synthesizing a type for a term ($\Gamma \vdash t \tpsynth T$)}
\label{fig:tp}
\end{figure}

\begin{figure}
  \[
  \begin{array}{lllllll}
    |x| & = & x &\ &
    |\abs{\lambda}{x}{t'}{t}| & = & \absu{\lambda}{x}{|t|} \\
    |t\ t'| & = & |t|\ |t'| &\ &
    |t\cdot T| & = & |t| \\
    |\absu{\Lambda}{x}{t'}{t}| & = & |t| &\ &
    |t\ \mhyph t'| & = & |t| \\
    |[t , t']| & = & |t| &\ &
    |t.1| & = & |t| \\
    |t.2| & = & |t| &\ &
    |\beta\{t\}| & = & |t|\\
    |\delta\ \mhyph\ t| & = & |t|&\ &
    |\rho\ t\ \mhyph\ t'| & = & |t'| \\
    |\phi\ t\ \mhyph\ t'\ \{t''\}| & = & |t''| &\ &
    |\chi\ T\ \mhyph\ t'| & = & |t'|
  \end{array}
  \]
  \caption{Erasure for annotated terms}
  \label{fig:eraser}
\end{figure}

@@

Defining \verb|Top| to be $\{\absu{\lambda}{x}{x} \simeq
\absu{\lambda}{x}{x}\}$, we may assign \verb|Top| to any closed term \verb|t|,
including non-normalizing ones. In our annotated syntax, we write
\texttt{\(\beta\)\{t\}}. Even without this, the presence of $\delta$ in
combination with $\phi$ allows us to type non-normalizing terms assuming an
erased argument $x$ of type $\{ \textit{tt} \simeq \textit{ff} \}$ for
Church-encoded booleans \textit{tt} and \textit{ff}. For example, $\delta\ \mhyph x$
has type \(\{\absu{\lambda}{x}{x} \simeq \absu{\lambda}{x}{x\ x}\}\), and with
$\phi$ we can use this to type $\Omega$ by changing the typed term
\(\texttt{id}\ \cdot\texttt{True}\ \texttt{id}\), where \verb|True| is
\(\abs{\forall}{X}{\star}{X \to X}\). But failure of normalization does not
impinge on Theorem~\ref{thm:consis}. Extensional Martin-L\"of type theory (MLTT)
is also non-normalizing, for a very similar reason, but fact does not contradict
its logical soundness~\cite{dybjer16}. In CDLE, the guarantees one gets about
the behavior of terms are expressed almost entirely in their types. If the types
are weak, then not much is guaranteed; but stronger types can guarantee
properties like normalization, as demonstrated by the following theorem:

>>>>

    \infer{\Gamma\vdash \beta\{t'\} \tpcheck \{ t \simeq t \}}{\Gamma\vdash \textit{FV}(t)\subseteq \textit{dom}(\Gamma)}  \\ \\    
    \infer{\Gamma\vdash \delta\ \mhyph\ t \tpcheck T}{\Gamma\vdash t\tpcheck \{ \absu{\lambda}{x}{\absu{\lambda}{y}{x}} \simeq \absu{\lambda}{x}{\absu{\lambda}{y}{y}}\}}  &
    \infer{\Gamma\vdash \rho\ t'\ \mhyph\ t \Leftrightarrow [t_2/x]T}
          {\Gamma\vdash t' \tpsynth \{ t_1 \simeq t_2 \} & \Gamma \vdash t \Leftrightarrow [t_1/x]T} \\ \\
    \infer{\Gamma\vdash \chi\ T\ \mhyph\ t \tpcheck T'}
          {\Gamma\vdash T\tpcheck \star & \Gamma\vdash t \tpcheck T & T \cong T'} &
    \infer{\Gamma\vdash \chi\ T\ \mhyph\ t \tpsynth T}
          {\Gamma\vdash T\tpcheck \star & \Gamma\vdash t \tpsynth T' & T \cong T'} \\ \\
    \infer{\Gamma\vdash \phi\ t\ \mhyph\ t'\{t''\} \Leftrightarrow T}
          {\Gamma\vdash t\tpsynth \{t'\simeq t''\} & \Gamma\vdash t' \Leftrightarrow T}  & \
  \end{array}
  \]
\caption{Rules for checking a term against a well-kinded type ($\Gamma \vdash t \tpcheck T$)
           and synthesizing a type for a term ($\Gamma \vdash t \tpsynth T$)}
\label{fig:tp}
\end{figure}

\begin{figure}
  \[
  \begin{array}{lllllll}
    |x| & = & x &\ &
    |\abs{\lambda}{x}{t'}{t}| & = & \absu{\lambda}{x}{|t|} \\
    |t\ t'| & = & |t|\ |t'| &\ &
    |t\cdot T| & = & |t| \\
    |\abs{\Lambda}{x}{t'}{t}| & = & |t| &\ &
    |t\ \mhyph t'| & = & |t| \\
    |[t , t']| & = & |t| &\ &
    |t.1| & = & |t| \\
    |t.2| & = & |t| &\ &
    |\beta\{t\}| & = & |t|\\
    |\delta\ \mhyph\ t| & = & |t|&\ &
    |\rho\ t\ \mhyph\ t'| & = & |t'| \\
    |\phi\ t\ \mhyph\ t'\ \{t''\}| & = & |t''| &\ &
    |\chi\ T\ \mhyph\ t'| & = & |t'|
  \end{array}
  \]
  \caption{Erasure for annotated terms}
  \label{fig:eraser}
\end{figure}

@@

Defining \verb|Top| to be $\{\absu{\lambda}{x}{x} \simeq
\absu{\lambda}{x}{x}\}$, we may assign \verb|Top| to any closed term \verb|t|,
including non-normalizing ones. In our annotated syntax, we write
\texttt{\(\beta\)\{t\}}. Even without this, the presence of $\delta$ in
combination with $\phi$ allows us to type non-normalizing terms assuming an
erased argument $x$ of type $\{ \textit{tt} \simeq \textit{ff} \}$ for
Church-encoded booleans \textit{tt} and \textit{ff}. For example, $\delta\ \mhyph x$
has type \(\{\absu{\lambda}{x}{x} \simeq \absu{\lambda}{x}{x\ x}\}\), and with
$\phi$ we can use this to type $\Omega$ by changing the typed term
\(\texttt{id}\ \cdot\texttt{True}\ \texttt{id}\), where \verb|True| is
\(\abs{\forall}{X}{\star}{X \to X}\). But failure of normalization does not
impinge on Theorem~\ref{thm:consis}. Extensional Martin-L\"of type theory (MLTT)
is also non-normalizing, for a very similar reason, but this fact does not contradict
its logical soundness~\cite{dybjer16}. In CDLE, the guarantees one gets about
the behavior of terms are expressed almost entirely in their types. If the types
are weak, then not much is guaranteed; but stronger types can guarantee
properties like normalization, as demonstrated by the following theorem:
